Data Pre-Processing

### 2021-05-04 AST-Framing-Ontario/Data Pre-Processesing.rmd
May 4th update

Load the following packages:
```{r}
library(tm)
library(readr)
library(tools)
library(dplyr)
library(tidytext)
library(topicmodels)
library(ggplot2)
library(wordcloud)
library(word2vec)
library(pdftools)
library(reshape2)
```

*Parametrize the document to run different corpus*

@@

In total, _X_ articles were accessed.

@@

```{r}
pdfLocation <- "Texts/School Boards"

@@

NAwords <- c("contact", "cancellations", "catholic", "waterloo", "board", "east", "transportation", "students", "student", "website", "tri", "can", "district", "will", "please", "may", "school", "https", "school", "schools", "walking", "cycling", "gta", "login", "and", "or", "the", "active", "asst", "stp")
corpus <- tm_map(corpus, removeWords, NAwords)
corpus <- tm_map(corpus, removeNumbers)

@@

```{r}

# Generate a random seed to use 
numIteration <- 10
r <- 1 
col <- 4 
betaList <- c("beta.0")
seeds <- sample.int(10000, numIteration-1)
avgtopics <- topics

for (seed1 in seeds){
  templda <- LDA(DTM, k, control = list(seed = seed1))
  temptopics <- tidy(templda, matrix = "beta")
  avgtopics <- left_join(avgtopics, temptopics , by = c("topic", "term"))
  colnames(avgtopics)[col] <- paste("beta.", r, sep = "")
  betaList <- append(betaList, paste("beta.", r, sep = ""))
  r <- r + 1
  col <- col + 1 
}
colnames(avgtopics)[3] <- "beta.0"

avgtopics$beta.mean <- rowMeans(avgtopics[betaList])

# The value below determines the number of top terms
avg_top_terms <- avgtopics %>% 
  group_by(topic) %>% 
  top_n(15, beta.mean) %>% 
  ungroup() %>% 
  arrange(topic, -beta.mean)
avg_top_terms

# Beta represents the probability of 'term' being generated in topic 'n' 
# Topics are generated by the LDA algorithm, it's up to the user to interpret the context of the topics

avg_top_terms %>% 
  mutate(term = reorder_within(term, beta.mean, topic)) %>%
  ggplot(aes(beta.mean, term, fill = factor(topic))) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~topic, scales = "free") + 
  scale_y_reordered() +
  scale_x_continuous(limits = c(0,0.06))

@@

# Word to Vector Test

@@

```{r}
library(text2vec)

# Create new data frame with the cleaned corpus 
cleanTextDF <- data.frame(text = sapply(corpus, as.character), stringsAsFactors = FALSE)

@@

Corpora will be cleaned for white spaces, punctuation, English stop words, and converted to lowercase. Any additional stop words will need to be removed manually as the user sees fit. 
```{r}
# Iterating through each subfolders 
m <- 1
for (subFolder in subFolders){

@@

  for (address in filePath)
{
doc <- readPDF(control = list(text = "-layout"))(elem = list(uri = address), language = "en")
test <- doc$content
# "file_path_sans_ext(basename())" from the tools package strips the base name for the file of interest and removed the extension at the end which as seen below allow for the replacement of .pdf for .txt 
# use of 'sep = ""' argument is important to prevent the addition of spaces, which is the default for the paste function 
newFile <- paste(txtLocation, "/", file_path_sans_ext(basename(address)), ".txt", sep = "")
write(test, file = newFile)
print( paste(file_path_sans_ext(basename(address)), ".txt Has been created", sep = ""))
# Creates uniquely id text store in 'Environment'
assign(paste("text", n, sep = ""), readChar(newFile, file.info(newFile)$size))
n <- n+1
  }

@@

temp_df_source <- DataframeSource(textDF)
tempCorpus <- VCorpus(temp_df_source)
tempCorpus <- tm_map(tempCorpus, stripWhitespace)
tempCorpus <- tm_map(tempCorpus, content_transformer(tolower))
tempCorpus <- tm_map(tempCorpus, removePunctuation)
tempCorpus <- tm_map(tempCorpus, removeWords, stopwords("english"))

@@

@@

}
rm(temp_df_source, tempCorpus, matrix1, matrix2, address, doc_id, fileBaseNames, filePath, i, m, n, newFile, pdfLocation, subFolder, test, txtLocation)
```

Uncomment and use the code below to further clean you corpuses. Insert stopwords as list of strings in 'NAwords'.  
```{r}
#NAwords <- c()
#corpus <- tm_map(corpus, removeWords, NAwords)
#corpus <- tm_map(corpus, removeNumbers)
```

Creating Term-Document Matrix and Document-Term Matrix with the fully cleaned corpora.
```{r}
# n - number of corpora
n <- 3
for (i in seq(1, n, 1))
{
tempCorpus <-  eval(parse(text = paste("corpus", toString(i), sep = "")))
TDM <- TermDocumentMatrix(tempCorpus)
DTM <- DocumentTermMatrix(tempCorpus)
matrix1 <- as.matrix(TDM)
matrix2 <- as.matrix(DTM)

@@

}
```

@@

```{r}
sort(NAwords)
```

>>>>

Load the following packages:
```{r}
library(tm)
library(readr)
library(tools)
library(dplyr)
library(tidytext)
library(topicmodels)
library(ggplot2)
library(wordcloud)
#library(word2vec)
library(pdftools)
library(reshape2)
library(text2vec)
```

*Parametrize the document to run different corpus* 
[We could parametrize specific stop words base on the specific files]

@@

In total, 233 articles were included from a list of 253. Reasons of exclusions for remaining 20 include unavailablity online, access only on print, and published in non-english.

@@

```{r}
pdfLocation <- "Texts/Academic Papers NoRef"

@@

NAwords <- c("school", "transportation", "can", "use", "active", "et", "al", "travel", "study", "elsevier", "na", "sciencedirect", "journal", "transport", "mode", "modes", "children", "the", "use", "n", "s", "schools", "also", "p", "b", "used", "number")
corpus <- tm_map(corpus, removeWords, NAwords)
corpus <- tm_map(corpus, removeNumbers)

@@

NOTE: RUNNING THE BLOCK BELOW WITH 'Academic Articles'AS INPUT WILL RESULT IN VERY LONG PROCESSING TIME

```{r}
# 
# # Generate a random seed to use 
# numIteration <- 10
# r <- 1 
# col <- 4 
# betaList <- c("beta.0")
# seeds <- sample.int(10000, numIteration-1)
# avgtopics <- topics
# 
# for (seed1 in seeds){
#   templda <- LDA(DTM, k, control = list(seed = seed1))
#   temptopics <- tidy(templda, matrix = "beta")
#   avgtopics <- left_join(avgtopics, temptopics , by = c("topic", "term"))
#   colnames(avgtopics)[col] <- paste("beta.", r, sep = "")
#   betaList <- append(betaList, paste("beta.", r, sep = ""))
#   r <- r + 1
#   col <- col + 1 
# }
# colnames(avgtopics)[3] <- "beta.0"
# 
# avgtopics$beta.mean <- rowMeans(avgtopics[betaList])
# 
# # The value below determines the number of top terms
# avg_top_terms <- avgtopics %>% 
#   group_by(topic) %>% 
#   top_n(15, beta.mean) %>% 
#   ungroup() %>% 
#   arrange(topic, -beta.mean)
# avg_top_terms
# 
# # Beta represents the probability of 'term' being generated in topic 'n' 
# # Topics are generated by the LDA algorithm, it's up to the user to interpret the context of the topics 
# 
# avg_top_terms %>% 
#   mutate(term = reorder_within(term, beta.mean, topic)) %>%
#   ggplot(aes(beta.mean, term, fill = factor(topic))) + 
#   geom_col(show.legend = FALSE) + 
#   facet_wrap(~topic, scales = "free") + 
#   scale_y_reordered() +
#   scale_x_continuous(limits = c(0,0.06))

@@

# GloVe Word Embeddings

@@

```{r}
# Create new data frame with the cleaned corpus 
cleanTextDF <- data.frame(text = sapply(corpus, as.character), stringsAsFactors = FALSE)

@@

Corpora will be cleaned for white spaces, punctuation, English stop words, and converted to lowercase. Any additional stop words will need to be removed manually as the user sees fit. 'corpusKey' is a data frame that indicates which file a corpus was derived from. 
```{r}
# Stop words for each of the corpora, or based on how your directory organizes files 
# Add stop words if you have more than 4 corpora 
NAwords2 <- c("login", "css", "sir", "waterloo", "niagara", "john", "school", "transportation", "york", "ottawa", "can", "tip", "london", "provides", "elgin", "find", "make", "active", "travel", "lord", "holy", "toronto", "-")
NAwords3 <- c("york", "lane", "school", "transportation", "region", "halton", "can", "use", "guelph", "kingston", "durham", "get")
NAwords4 <- c("active", "and", "asst", "board", "can", "cancellations", "catholic", "contact", "cycling", "district", "east", "gta", "https", "login", "may", "or", "please", "school", "schools", "stp", "student", "students", "the", "transportation", "tri", "walking", "waterloo", "website", "will")
NAwords1 <- c("school", "transportation", "can", "use", "active", "et", "al", "travel", "study", "elsevier", "na", "sciencedirect", "journal", "transport", "mode", "modes", "children", "the", "use", "n", "s", "schools", "also", "p", "b", "used", "number", "data", "research", "variable", "variables", "author", "authors", "researcher", "researchers", "project", "intrument", "may")

corpusNum <- c()

# Iterating through each subfolders 
m <- 1
for (subFolder in subFolders){

@@

# "file_path_sans_ext(basename())" from the tools package strips the base name for the file of interest and removed the extension at the end which as seen below allow for the replacement of .pdf for .txt 
# use of 'sep = ""' argument is important to prevent the addition of spaces, which is the default for the paste function 
newFile <- paste(txtLocation, "/", file_path_sans_ext(basename(address)), ".txt", sep = "")
write(test, file = newFile)

####print(paste(file_path_sans_ext(basename(address)), ".txt Has been created", sep = ""))

# Creates uniquely id text store in 'Environment'
assign(paste("text", n, sep = ""), readChar(newFile, file.info(newFile)$size))
n <- n+1
  }

@@

temp_df_source <- DataframeSource(textDF)
tempCorpus <- VCorpus(temp_df_source)

# corpus cleaning 
tempCorpus <- tm_map(tempCorpus, stripWhitespace)
tempCorpus <- tm_map(tempCorpus, content_transformer(tolower))
tempCorpus <- tm_map(tempCorpus, removePunctuation)
tempCorpus <- tm_map(tempCorpus, removeWords, stopwords("english"))
tempCorpus <- tm_map(tempCorpus, removeWords, eval(parse(text = paste("NAwords", m, sep = ""))))
tempCorpus <- tm_map(tempCorpus, removeNumbers)

assign(paste("cleanTextDF", m, sep = ""), data.frame(text = sapply(tempCorpus, as.character), stringsAsFactors = FALSE))

@@

corpusNum <- c(corpusNum, paste(m))

@@

}

corpusKey <- data.frame(corpusNum, subFolders)

rm(temp_df_source, tempCorpus, address, doc_id, fileBaseNames, filePath, i, m, n, newFile, pdfLocation, subFolder, test, txtLocation)
```

Creating Term-Document Matrix and Document-Term Matrix with the fully cleaned corpora.
```{r}
# n - number of corpora
n <- 4
for (i in seq(1, n, 1))
{
tempCorpus <-  eval(parse(text = paste("corpus", i, sep = "")))
TDM <- TermDocumentMatrix(tempCorpus)
DTM <- DocumentTermMatrix(tempCorpus)
matrix1 <- as.matrix(TDM)
matrix2 <- as.matrix(DTM)

@@

}
rm(matrix1, matrix2)
```

# Vocabulary

```{r}
# n - number of corpora
n <- 4

for (i in seq(1, n, 1))
{
  tempDF <- eval(parse(text = paste("cleanTextDF", i, sep = "")))
  it_train = itoken(tempDF$text, progressbar = FALSE)
  assign(paste("vocab", i, sep= ""), create_vocabulary(it_train))
  tempVocab <- create_vocabulary(it_train)
  assign(paste("vocab", i, sep= ""), prune_vocabulary(tempVocab, term_count_min = 5))
}

rm(tempVocab)
```

@@

```{r}
#eval(parse(text = paste("NAwords", 1, sep = "")))
NAwords1
